---
title: "US.Aus.MHWradius1.v2.1"
output: html_document
---

This code links marine heatwave metrics to each coastal community. Through this script, we modified code from Kurt C. Heim. 

#Load in all packages, functions, and datsets needed for this code to run
```{r}
source("/Users/sallydowd/Documents/GitHub/Nye-research-technicians/MHW.MCS/Code/ToLoad.R")
```

#MHW/MCS analysis for mainland US communities 

##Link MHW statistics with each mainland US community for 2012-2016
```{r}
###Cumulative intensity
US_main_MHW <- read_csv("/Users/sallydowd/Documents/GitHub/Nye-research-technicians/MHW.MCS/Data/MHW.block.main.V2.1.csv")
US_main_cumi_MHW <-aggregate(total_icum ~ lat_lon + lat + lon, US_main_MHW[between(US_main_MHW$year, 2012, 2016),], FUN = mean) ####Get average of total cumulative intensity for each set of coordinates between 2012-2016
US_main_MHW_geom = st_as_sf(US_main_cumi_MHW, coords=c("lon","lat"))
US_main_cumi_MHW_ld <- metrics_df(only_US_main_orig, radius1_US_main, US_main_MHW_geom, 5) ####Find intersection of 111 km around each community with MHW coordinates, take average of total cumulative intensity (averaged over 2012-2016) over these points 
US_main_cumi_MHW_ld <- US_main_cumi_MHW_ld %>% rename("cumi_N"= "input_N", "cumi_mean"= "input_mean", "cumi_var"= "input_var", "cumi_max"= "input_max", "cumi_min"= "input_min")
US_main_cumi_MHW_ld <- US_main_cumi_MHW_ld[!is.na(US_main_cumi_MHW_ld$cumi_mean),] 
write.csv(US_main_cumi_MHW_ld, "/Users/sallydowd/Documents/GitHub/Nye-research-technicians/MHW.MCS/Data/US.main.cumi.2016.V2.1.csv")

###Duration 
US_main_dur_MHW <-aggregate(duration ~ lat_lon + lat + lon, US_main_MHW[between(US_main_MHW$year, 2012, 2016),], FUN = mean)
US_main_dur_geom = st_as_sf(US_main_dur_MHW, coords=c("lon","lat"))
US_main_dur_MHW_ld <- metrics_df(only_US_main_orig, radius1_US_main, US_main_dur_geom, 5) 
US_main_dur_MHW_ld <- US_main_dur_MHW_ld %>% rename("dur_N"= "input_N", "dur_mean"= "input_mean", "dur_var"= "input_var", "dur_max"= "input_max", "dur_min"= "input_min")
US_main_dur_MHW_ld <- US_main_dur_MHW_ld[!is.na(US_main_dur_MHW_ld$dur_mean),] 
write.csv(US_main_dur_MHW_ld, "/Users/sallydowd/Documents/GitHub/Nye-research-technicians/MHW.MCS/Data/US.main.dur.2016.V2.1.csv")
```

##Link MHW statistics with each HI community for 2012-2016 
```{r}
###Cumulative intensity
US_HI_MHW <- read_csv("/Users/sallydowd/Documents/GitHub/Nye-research-technicians/MHW.MCS/Data/MHW.block.HI.V2.1.csv")
US_HI_cumi_MHW <-aggregate(total_icum ~ lat_lon + lat + lon, US_HI_MHW[between(US_HI_MHW$year, 2012, 2016),], FUN = mean) 
US_HI_MHW_geom = st_as_sf(US_HI_cumi_MHW, coords=c("lon","lat"))
US_HI_cumi_MHW_ld <- metrics_df(only_US_HI_orig, radius1_HI, US_HI_MHW_geom, 5) 
US_HI_cumi_MHW_ld <- US_HI_cumi_MHW_ld %>% rename("cumi_N"= "input_N", "cumi_mean"= "input_mean", "cumi_var"= "input_var", "cumi_max"= "input_max", "cumi_min"= "input_min")
US_HI_cumi_MHW_ld <- US_HI_cumi_MHW_ld[!is.na(US_HI_cumi_MHW_ld$cumi_mean),] 
write.csv(US_HI_cumi_MHW_ld, "/Users/sallydowd/Documents/GitHub/Nye-research-technicians/MHW.MCS/Data/US.HI.cumi.2016.V2.1.csv")

###Duration
US_HI_dur_MHW <-aggregate(duration ~ lat_lon + lat + lon, US_HI_MHW[between(US_HI_MHW$year, 2012, 2016),], FUN = mean)
US_HI_dur_geom = st_as_sf(US_HI_dur_MHW, coords=c("lon","lat"))
US_HI_dur_MHW_ld <- metrics_df(only_US_HI_orig, radius1_HI, US_HI_dur_geom, 5) 
US_HI_dur_MHW_ld <- US_HI_dur_MHW_ld %>% rename("dur_N"= "input_N", "dur_mean"= "input_mean", "dur_var"= "input_var", "dur_max"= "input_max", "dur_min"= "input_min")
US_HI_dur_MHW_ld <- US_HI_dur_MHW_ld[!is.na(US_HI_dur_MHW_ld$dur_mean),] 
write.csv(US_HI_dur_MHW_ld, "/Users/sallydowd/Documents/GitHub/Nye-research-technicians/MHW.MCS/Data/US.HI.dur.2016.V2.1.csv")
```

##Link MHW statistics with each AK community for 2012-2016
```{r}
###Cumulative intensity
US_AK_MHW <- read_csv("/Users/sallydowd/Documents/GitHub/Nye-research-technicians/MHW.MCS/Data/MHW.block.AK.V2.1.csv")
US_AK_cumi_MHW <-aggregate(total_icum ~ lat_lon + lat + lon, US_AK_MHW[between(US_AK_MHW$year, 2012, 2016),], FUN = mean) 
US_AK_MHW_geom = st_as_sf(US_AK_cumi_MHW, coords=c("lon","lat"))
US_AK_cumi_MHW_ld <- metrics_df(only_US_AK_orig, radius1_AK, US_AK_MHW_geom, 5) 

US_AK_cumi_MHW_ld <- US_AK_cumi_MHW_ld %>% rename("cumi_N"= "input_N", "cumi_mean"= "input_mean", "cumi_var"= "input_var", "cumi_max"= "input_max", "cumi_min"= "input_min")
US_AK_cumi_MHW_ld <- US_AK_cumi_MHW_ld[!is.na(US_AK_cumi_MHW_ld$cumi_mean),] 
write.csv(US_AK_cumi_MHW_ld, "/Users/sallydowd/Documents/GitHub/Nye-research-technicians/MHW.MCS/Data/US.AK.cumi.2016.V2.1.csv")

###Duration
US_AK_dur_MHW <-aggregate(duration ~ lat_lon + lat + lon, US_AK_MHW[between(US_AK_MHW$year, 2012, 2016),], FUN = mean)
US_AK_dur_geom = st_as_sf(US_AK_dur_MHW, coords=c("lon","lat"))
US_AK_dur_MHW_ld <- metrics_df(only_US_AK_orig, radius1_AK, US_AK_dur_geom, 5) 
US_AK_dur_MHW_ld <- US_AK_dur_MHW_ld %>% rename("dur_N"= "input_N", "dur_mean"= "input_mean", "dur_var"= "input_var", "dur_max"= "input_max", "dur_min"= "input_min")
US_AK_dur_MHW_ld <- US_AK_dur_MHW_ld[!is.na(US_AK_dur_MHW_ld$dur_mean),] 
write.csv(US_AK_dur_MHW_ld, "/Users/sallydowd/Documents/GitHub/Nye-research-technicians/MHW.MCS/Data/US.AK.dur.2016.V2.1.csv")
```


##Edit dataset

Combine cumulative intensity and duration datasets for the US and filter the dataset to only retain communities present in indicator dataset (US.indicator.data.final.csv). This removes communities with no MHW statistics, those in the Ches Bay, those with an NA for at least one column for indicator data)
```{r}
###Part 1: Combine cumulative intensity datasets: 
US_main_cumi_MHW_ld <- read_csv("/Users/sallydowd/Documents/GitHub/Nye-research-technicians/MHW.MCS/Data/US.main.cumi.2016.V2.1.csv")
US_AK_cumi_MHW_ld <- read_csv("/Users/sallydowd/Documents/GitHub/Nye-research-technicians/MHW.MCS/Data/US.AK.cumi.2016.V2.1.csv") 
US_HI_cumi_MHW_ld <- read_csv("/Users/sallydowd/Documents/GitHub/Nye-research-technicians/MHW.MCS/Data/US.HI.cumi.2016.V2.1.csv")
US_total_cumi_MHW_ld <- rbind(US_main_cumi_MHW_ld, US_AK_cumi_MHW_ld, US_HI_cumi_MHW_ld)
#write.csv(US_total_cumi_MHW_ld, "/Users/sallydowd/Documents/GitHub/Nye-research-technicians/MHW.MCS/Data/US.total.cumi.2016.V2.1.csv")

###Part 2: Step 3) Filter US cumulative intensity dataset to match communities in the Indicator_data_final (steps completed already: 1) Filtered Indicator_data based off communities in US_total_cumi_MHW_ld so communities that had MHW stats 2) Removed communities in Ches Bay 3) Removed incomplete cases based off missing parameters for indicators)
Indicator_data_final <- read.csv(file= "/Users/sallydowd/Documents/GitHub/Nye-research-technicians/MHW.MCS/Data/US.indicator.data.final.csv") ####these are the final coastal communities for cumulative intensity so we need to make sure these are the only ones we use in MHW analysis!!

US_total_cumi_MHW_ld$Community.Name <- paste(US_total_cumi_MHW_ld$Community.Name, US_total_cumi_MHW_ld$State, sep = ", ")
US_coastal_communities_indicator <- Indicator_data_final$Community.Name 
US_total_cumi_MHW_ld_final <- US_total_cumi_MHW_ld %>% filter(Community.Name %in% US_coastal_communities_indicator) %>% arrange(Community.Name) 
#write.csv(US_total_cumi_MHW_ld_final, "/Users/sallydowd/Documents/GitHub/Nye-research-technicians/MHW.MCS/Data/US.total.cumi.2016.final.V2.1.csv")

###Part 3: Combine duration datasets and filter for communities present in indicator analysis
US_main_dur_MHW_ld <- read_csv("/Users/sallydowd/Documents/GitHub/Nye-research-technicians/MHW.MCS/Data/US.main.dur.2016.V2.1.csv")
US_AK_dur_MHW_ld <- read_csv("/Users/sallydowd/Documents/GitHub/Nye-research-technicians/MHW.MCS/Data/US.AK.dur.2016.V2.1.csv") 
US_HI_dur_MHW_ld <- read_csv("/Users/sallydowd/Documents/GitHub/Nye-research-technicians/MHW.MCS/Data/US.HI.dur.2016.V2.1.csv")

US_total_dur_MHW_ld <- rbind(US_main_dur_MHW_ld, US_AK_dur_MHW_ld, US_HI_dur_MHW_ld)
US_total_dur_MHW_ld$Community.Name <- paste(US_total_dur_MHW_ld$Community.Name, US_total_dur_MHW_ld$State, sep = ", ")
US_total_dur_MHW_ld_final <- US_total_dur_MHW_ld %>% filter(Community.Name %in% US_coastal_communities_indicator)
#write.csv(US_total_dur_MHW_ld_final, "/Users/sallydowd/Documents/GitHub/Nye-research-technicians/MHW.MCS/Data/US.total.dur.2016.final.V2.1.csv")
```

#MHW/MCS analysis for Australia LGAs  

##Link MHW statistics with each Australia LGA for 2012-2016
```{r}
###Cumulative intensity
Australia_MHW <- read_csv("/Users/sallydowd/Documents/GitHub/Nye-research-technicians/MHW.MCS/Data/MHW.block.Aus.V2.1.csv")
####A) Cumulative intensity
Australia_cumi_MHW <-aggregate(total_icum ~ lat_lon + lat + lon, Australia_MHW[between(Australia_MHW$year, 2012, 2016),], FUN = mean)
Australia_MHW_geom = st_as_sf(Australia_cumi_MHW, coords=c("lon","lat"))
Australia_cumi_MHW_ld <- metrics_df(only_Aus_cs_orig, radius1_Aus, Australia_MHW_geom, 11)
Australia_cumi_MHW_ld <- Australia_cumi_MHW_ld %>% rename("cumi_N"= "input_N", "cumi_mean"= "input_mean", "cumi_var"= "input_var", "cumi_max"= "input_max", "cumi_min"= "input_min", "LGA"= "LGA_NAME16")
Australia_cumi_MHW_ld <- Australia_cumi_MHW_ld[!is.na(Australia_cumi_MHW_ld$cumi_mean),] 
write.csv(Australia_cumi_MHW_ld, "/Users/sallydowd/Documents/GitHub/Nye-research-technicians/MHW.MCS/Data/Australia.cumi.2016.V2.1.csv")

###Duration
Australia_dur_MHW <-aggregate(duration ~ lat_lon + lat + lon, Australia_MHW[between(Australia_MHW$year, 2012, 2016),], FUN = mean)
Australia_MHW_geom_dur = st_as_sf(Australia_dur_MHW, coords=c("lon","lat"))
Australia_dur_MHW_ld <- metrics_df(only_Aus_cs_orig, radius1_Aus, Australia_MHW_geom_dur, 11)
Australia_dur_MHW_ld <- Australia_dur_MHW_ld %>% rename("dur_N"= "input_N", "dur_mean"= "input_mean", "dur_var"= "input_var", "dur_max"= "input_max", "dur_min"= "input_min", "LGA"= "LGA_NAME16")
Australia_dur_MHW_ld <- Australia_dur_MHW_ld[!is.na(Australia_dur_MHW_ld$dur_mean),] 
write.csv(Australia_dur_MHW_ld, "/Users/sallydowd/Documents/GitHub/Nye-research-technicians/MHW.MCS/Data/Australia.dur.2016.V2.1.csv")
```
